A fully labelled proof system for intuitionistic modal logics

### 2021-01-14 FullyLabelled/JLC/introduction.tex
margins fixed
typo

Labelled sequent calculi for logics with
standard Kripke semantics attach to every formla $\fm A$ a label
$\lb x$, witten as $\labels xA$, and additionally use
\emph{relational atoms} of the form $\accs xy$ where 
the binary relation symbol $\rel$
%These calculi work best for logics with
%standard Kripke semantics, as in this case the relation $\rel$ 
is used to encode the accessibility relation in the Kripke models, such that
frame conditions corresponding to the desired logic can be directly
encoded as inference rules.
%\todo{Reviewer 1 complained about the ``best''}
Prominent examples are classical modal
logics and intuitionistic propositional logic, e.g., the frame
condition of transitivity ($\forall \lb {xyz}.\:\accs xy\AND \accs yz\IMP\accs xz$) can be translated into the
inference rule
\begin{equation}
\label{eq:Rtrans}
\vlinf{\rn{trans}}{}{\B, \accs xy, \accs yz, \Left \SEQ \Right}{
	\B, \accs xy, \accs yz, \accs xz, \Left \SEQ \Right}
\end{equation}
where $\B$ stands for a set of relational atoms, and $\Gamma$ and
$\Delta$ for multi-sets of labelled formulas~\cite{negri:jpl2005}.

@@

%\marianela{The reviewer suggested to change the above sentence for the next one:
%	\emph{"As done earlier in the literature  on intuitionistic modal logic (cf....) we consider here
%	to enrich"} since Maffezioli, Naibo and Negri have already done this before and the next one:
%	D. Garg, V. Genovese, and S. Negri, Countermodels from sequent calculi in multi-modal	logics. LICS 2012, IEEE Computer Society, pp. 315–324, 2012. (ALREADY ADDED ON REFERENCES FILE TO USE IT IF WE CHANGE THE SENTENCE: \cite{garg2012})
%	What do you think about changing the sentence? We have already mentioned Maffezioli, Naibo and Negri's paper a few times...
%	}
\todo{add footnote}\sonia{added the reference in the text directly}

@@

A system similar to ours is presented
in~\cite{maffezioli:etal:synthese13}, for intuitionistic bimodal
epistemic logic. That system is also a multi-conclusion labelled system with
more than one relation. In fact, the logical rules are essentially the
same as ours, but we use a more general setting and provide deductive proof
systems for a larger class of intuitionistic modal logics, together
with a syntactic cut elimination procedure.
%\review{More specifically, it is not clear what the authors mean by  "internal cut elimination procedure" on the next paragraph.
%	At the beginning of the paper "internal calculi" are identified as those calculi, such as hypersequent
%	or nested calculi, that do not have explicit accessibility relations. But then in which sense it the
%	cut elimination "internal"? Is it because there are no cuts on relational atoms? This is the case
%	also for the mentioned MNN paper.  Please clarify.}
\sonia{OK}

@@

This paper is organized as follows, In the next section (Section~\ref{sec:intmod}) we recall the standard syntax and semantics of intuitionistic modal logics. Then, in Section~\ref{sec:system} we present our system for the intuitionistic modal logic $\IK$. In Sections~\ref{sec:completeness} and~\ref{sec:soundness}, we show soundness and completeness of the system with cut. The cut elimination theorem, proved in Section~\ref{sec:cut-elim}, then entails soundness and completeness for the cut-free system. Finally, in Section~\ref{sec:ext} we discuss the possible extension to the system to capture intuitionistic Scott-Lemmon modal logics.

>>>>

Labelled sequent calculi for logics with
standard Kripke semantics attach to every formla $\fm A$ a label
$\lb x$, written as $\labels xA$, and additionally use
\emph{relational atoms} of the form $\accs xy$ where 
the binary relation symbol $\rel$
%These calculi work best for logics with
%standard Kripke semantics, as in this case the relation $\rel$ 
is used to encode the accessibility relation in the Kripke models, such that
frame conditions corresponding to the desired logic can be directly
encoded as inference rules.
%\todo{Reviewer 1 complained about the ``best''}
Prominent examples are classical modal
logics and intuitionistic propositional logic, e.g., the frame
condition of transitivity ($\forall \lb {xyz}.\:\accs xy\AND \accs yz\IMP\accs xz$) can be translated into the
inference rule
\begin{equation}
\label{eq:Rtrans}
\vlinf{\rn{trans}}{}{\B, \accs xy, \accs yz, \Left \SEQ \Right}{
	\B, \accs xy, \accs yz, \accs xz, \Left \SEQ \Right}
\end{equation}
where $\B$ stands for a set of relational atoms, and $\Gamma$ and
$\Delta$ for multi-sets of labelled formulas~\cite{negri:jpl2005}.

@@

@@

A system similar to ours is presented
in~\cite{maffezioli:etal:synthese13}, for intuitionistic bimodal
epistemic logic. That system is also a multi-conclusion labelled system with
more than one relation. In fact, the logical rules are essentially the
same as ours, but we use a more general setting and provide deductive proof
systems for a larger class of intuitionistic modal logics, together
with a syntactic cut elimination procedure.
%\review{More specifically, it is not clear what the authors mean by  "internal cut elimination procedure" on the next paragraph.
%	At the beginning of the paper "internal calculi" are identified as those calculi, such as hypersequent
%	or nested calculi, that do not have explicit accessibility relations. But then in which sense it the
%	cut elimination "internal"? Is it because there are no cuts on relational atoms? This is the case
%	also for the mentioned MNN paper.  Please clarify.}

@@

This paper is organized as follows. In the next section (Section~\ref{sec:intmod}) we recall the standard syntax and semantics of intuitionistic modal logics. Then, in Section~\ref{sec:system} we present our system for the intuitionistic modal logic $\IK$. In Sections~\ref{sec:completeness} and~\ref{sec:soundness}, we show soundness and completeness of the system with cut. The cut elimination theorem, proved in Section~\ref{sec:cut-elim}, then entails soundness and completeness for the cut-free system. Finally, in Section~\ref{sec:ext} we discuss the possible extension to the system to capture intuitionistic Scott-Lemmon modal logics.
